---
title: "Exercise 11"
output: statsr:::statswithr_lab
author: Jacob John
---

<div id="instructions">
Complete all **Exercises**, and submit answers to **VtopBeta**
</div>


##Linear Regression

The steps to create the relationship is −

* Carry out the experiment of gathering a sample of observed values of height and corresponding weight.

* Create a relationship model using the _lm()_ functions in R.

* Find the coefficients from the model created and create the mathematical equation using these

* Get a summary of the relationship model to know the average error in prediction. Also called residuals.

* To predict the weight of new persons, use the _predict()_ function in R.

###Create Relationship Model
```{r}
# The predictor variable
x <- c(151, 174, 138, 186, 128, 136, 179, 163, 152, 131)

# The response variable
y <- c(63, 81, 56, 91, 47, 57, 76, 72, 62, 48)

# Apply the lm() function.
relation <- lm(y~x)
relation

```

###Get the Summary of the Relationship
```{r}
summary(relation)
```

###Predict
```{r}
# Find y for an x of 170.
a <- data.frame(x=170)
result <- predict(relation,a)
result
```

###Plots
```{r}
# Plot the chart.
plot(x,y,
     col = "blue",
     main = "Linear Regression",
     abline(lm(y~x)),
     cex = 1.3,pch = 16)

plot(lm(y~x))
```