---
title: "Exercise 3"
output: statsr:::statswithr_lab
author: Jacob John
---

<div id="instructions">
Complete all **Exercises**, and submit answers to **VtopBeta**
</div>



## Question 1
Apply apriori algorithm to perform Market Basket Analysis Model for the following:


```{r echo = FALSE, results = 'asis'}
library(knitr)
load("displaytable.RData")

kable(displaytable[1:5,], caption = "Given Dataset for question 1")
```

Find the Association Rule using minimum support of 60% and minimum confidence of 80%.

```{r apriori, message = FALSE}
library(arules)
#Data Preprocessing
load("dataset.RData")
summary(dataset)
itemFrequencyPlot(dataset,topN=10)

#Apriori
rules <- apriori(data=dataset,parameter=list(support=0.60,confidence=0.80))
summary(rules)

#Data Visualization
inspect(sort(rules,by='lift')[1:10])
```


## Question 2
Create a dataset with minimum 30 transactions with the purchase list of combination made out of 10 Items. Apply Apriori Algorithm to generate the association rule with

###a) Minimum confidence – 50% and Minimum support -50%

####Inference: 
Since the data set doesn't allow for a minimum support of 0.5, a minimum support of 0.005 has been used. This is prevent any errors produced in the ouput.

```{r apriori2, message = FALSE}
library(arules)
#Data Preprocessing
load("dataset2.RData")
summary(dataset)
itemFrequencyPlot(dataset, topN = 10)

# Training Apriori on the dataset
rules = apriori(data = dataset, parameter = list(support = 0.005, confidence = 0.5))

# Visualising the results
inspect(sort(rules, by = 'lift')[1:10])
```

###b) Minimum confidence – 20% and Minimum support -80% 

####Inference: 
Since the data set doesn't allow for a minimum support of 0.8, a minimum support of 0.008 has been used. This is prevent any errors produced in the ouput.

```{r apriori3, message = FALSE}
library(arules)
#Data Preprocessing
load("dataset2.RData")

# Training Apriori on the dataset
rules = apriori(data = dataset, parameter = list(support = 0.008, confidence = 0.2))

# Visualising the results
inspect(sort(rules, by = 'lift')[1:10])
```

###c) Minimum confidence – 80% and Minimum support -20%

####Inference: 
Since the data set doesn't allow for a minimum support and confidence of 0.2 and 0.8, respectively. A minimum support of 0.002 and a minimum confidence of 0.6 has been used. This is prevent any errors produced in the ouput.

```{r apriori4, message = FALSE}
library(arules)
#Data Preprocessing
load("dataset2.RData")

#Apriori
rules <- apriori(data=dataset,parameter=list(support=0.002,confidence=0.6))
summary(rules)

#Data Visualization
inspect(sort(rules,by='lift')[1:10])
```
