---
title: "Exercise 8"
output: statsr:::statswithr_lab
author: Jacob John
---

<div id="instructions">
Complete all **Exercises**, and submit answers to **VtopBeta**
</div>


##Introduction

*Note:* Install BIRCH package given in the "Birch" subfolder before you start this exercise.

###Working

This function creates a CF-Tree not unlike that described in Zhang et al. (1997), and used in Harrington and Salibian-Barrera (2010). A complete explanation of this package is given in Harrington and Salibian-Barrera (2008).


##Datasets

```{r data, message = FALSE}
## Create a data set
library(birch)
library(MASS)
set.seed(1234) 
x <- mvrnorm(1e5, mu=rep(0,5), Sigma=diag(1,5))
x <- rbind(x, mvrnorm(1e5, mu=rep(10,5), Sigma=diag(0.1,5)+0.9))
```

##Creating a BIRCH object
```{r object, message = FALSE}
## Creating object
birchObj <- birch(x, 5, keeptree = TRUE)
birchObj
```

##Leaving a tree in memory
```{r memory, message = FALSE}
## Leaving a tree in memory
birch.addToTree(x, birchObj)

birchObj <- birch.getTree(birchObj)
birchObj

## And don't forget to killTree...
birch.killTree(birchObj)
```
