---
title: "Exercise 10"
output: statsr:::statswithr_lab
author: Jacob John
---

<div id="instructions">
Complete all **Exercises**, and submit answers to **VtopBeta**
</div>


##Introduction

##Datasets

Using the iris dataset to compare the various Distance Measures found in Clustering – Euclidean distance, Manhattan Distance, Jaccard coefficient, Cosine distance and edit.

```{r echo = FALSE, results = 'asis'}
library(knitr)
library(ggplot2)
kable(iris[1:5,3:5], caption = "Iris dataset for finding distance matrices")
```


##Dist function

The *dist* function computes and returns the distance matrix computed by using the specified distance measure to compute the distances between the rows of a data matrix.

```{r complete, message = FALSE}

# Euclidean Distance
euclid_dist <- as.matrix(dist(iris[, 3:4], 
                              method = "euclidean", 
                              upper = TRUE, 
                              diag = TRUE))
euclid_dist[1:10,1:7]

# Manhattan Distance
man_dist <- as.matrix(dist(iris[, 3:4], 
                              method = "manhattan", 
                              upper = TRUE, 
                              diag = TRUE))
man_dist[1:10,1:10]

```


##Distance function
The *distance()* function implemented in philentropy is able to compute 46 different distances/similarities between probability density functions (see *?philentropy::distance* for details).

```{r philentropy, message = FALSE}
library(philentropy)

# Cosine distance
cos_dist <- as.matrix(distance(iris[,3:4], 
                               method = "cosine"))
cos_dist[1:6,1:7]

# Jaccard similarity
jac_sim <- as.matrix(distance(iris[,3:4], 
                               method = "jaccard"))
jac_sim[1:6,1:6]

```


##Approximate String Distances

Compute the approximate string distance between character vectors. The distance is a generalized Levenshtein (edit) distance, giving the minimal possibly weighted number of insertions, deletions and substitutions needed to transform one string into another.

```{r edit, message = FALSE}
#Edit Distance
edit_dist <- as.matrix(adist(iris[,5])) #y = NULL (default) indicating taking x as y
edit_dist[48:60,1:12]
```
