---
title: "Exercise 6"
output: statsr:::statswithr_lab
author: Jacob John
---

<div id="instructions">
Complete all **Exercises**, and submit answers to **VtopBeta**
</div>


##Loading the dataset
```{r echo = FALSE, results = 'asis'}
library(caret)
library(knitr)
library(mlbench)
kable(iris[1:5,1:5], caption = "Iris dataset for clustering")
ggplot(iris, aes(Petal.Length, Petal.Width, color = Species)) + geom_point()
```


##K-means

Since the initial cluster assignments are random, let us set the seed to ensure reproducibility.

```{r random, message = FALSE}
set.seed(20)
irisCluster <- kmeans(iris[,3:4],3,nstart=20)
irisCluster
```


Based on the above results, three species dominate the clusters. We then ask the algorithm to group the data into 3 clusters. 
```{r table, message = FALSE}
#Comparing clusters with the species
table(irisCluster$cluster, iris$Species)
irisCluster$cluster <- as.factor(irisCluster$cluster)
ggplot(iris, aes(Petal.Length, Petal.Width, color = irisCluster$cluster)) + geom_point()
```


##Inference
The data belonging to the setosa species got grouped into cluster 3, versicolor into cluster 2, and virginica into cluster 1.