---
title: "Exercise 2"
output: statsr:::statswithr_lab
author: Jacob John
---

<div id="instructions">
Complete all **Exercises**, and submit answers to **VtopBeta**
</div>



##Create Dataset
1) Create a csv file with at least 20 entries on the following attributes:<br />
Name Nominal<br />
Age 1…100<br />
Experience 1…15<br />

2) Remove values of Experience and Fill with "NA".

3) Replace few entries of Age with >100 values.

4) Print the dataset using R


```{r display, message = FALSE}
displaytable<-readRDS("exercisetwo.rds")
displaytable
```

##Missing data

1) Display the content of the dataset by removing the missing value tuples.

```{r noNA, message = FALSE}
exercise<-readRDS("exercisetwo.rds")
exercise[complete.cases(exercise),]
```

2) Calculate the mean of the attribute without considering the missing values.

```{r mean, message = FALSE}
data_mean <- mean(exercise$Age,na.rm=TRUE)
data_mean
```

3) Display the content of the dataset by replacing the 'NA' with mean of the attribute.

```{r display_content, message = FALSE}
exercisedata<-readRDS("exercisetwo.rds")
exercisedata$Experience[is.na(exercisedata$Experience)] <- mean(exercisedata$Experience,na.rm = TRUE)
exercisedata
```

##Outlier Detection

###Method 1

4) Replace the age value above 100 with "NA" and print the current dataset.

```{r replaceNA, message = FALSE}
exercisedata$Age[exercisedata$Age>100] <- NA
exercisedata
```

5) Replace the "NA" in age with Mean of age.
```{r replaceNAmean, message = FALSE}
exercisedata$Age[is.na(exercisedata$Age)] <- data_mean
exercisedata
```

###Method 2

6) Use the same dataset and change the value of age to lie outside the range of 100.
Apply box plot and scatter plot to find the outliers.

```{r boxplot, message = FALSE}
exercisedata$Age[is.na(exercisedata$Age)==data_mean] <- 145
exercisedata
boxplot(exercisedata$Age)
plot(exercisedata)
plot(exercisedata$Age)
```

##Redundancy Check
7) Repeat the experience column as new attribute Exp. Find the correlation between age, Exp & experience.

```{r redundancy, message = FALSE}
exercisedata$Exp <- exercisedata$Experience
exercisedata
corf <- cor(data.frame(exercisedata$Age,exercisedata$Experience,exercisedata$Exp))
corf
```
